Precise predictions for same-sign W-bosons scattering at the LHC

### 2018-02-15 Draft/shower.tex
further edits in shower.tex

\begin{table}[h!]
    \centering
    \begin{tabular}{c|r@{ $\pm$ }l}
      Code  &  \multicolumn{2}{c}{$\sigma[\rm{fb}]$}  \\
        \hline\hline
        {\sc MG5\_aMC}+{\sc Pythia8}&  $1.450 \, (1.368)$ & $0.004$  \\
        {\sc MG5\_aMC}+{\sc Herwig++}&  $1.445 \, (1.363)$ & $0.004$  \\
        {\sc POWHEG}  &  $1.3642$ & $0.0004$  \\
        {\sc VBFNLO}+{\sc Herwig7-Dipole} &  $1.3389$ & $0.0006$  \\
        {\sc VBFNLO}+{\sc Herwig7-Default} &  $1.3067$ & $0.0006$  \\
        \hline
        {\sc MG5\_aMC}+{\sc Pythia8} (LO)&  $1.352 \, (1.275)$ & $0.003$  \\
        {\sc MG5\_aMC}+{\sc Herwig++} (LO)&  $1.343 \, (1.267)$ & $ 0.003$  \\
        {\sc PHANTOM}+{\sc Pythia8} &  $1.235\phantom{0} $ & $0.001$  \\
        {\sc PHANTOM}+{\sc Herwig++} &  $1.260\phantom{0} $ & $0.001$  \\
        {\sc VBFNLO}+{\sc Herwig7} (LO) &  $1.3001$ & $0.0002$  \\
    \end{tabular}
    \caption{\label{tab:PSrates} Rates at NLO-QCD (LO-QCD) accuracy matched to parton shower within VBS cuts obtained with the different codes used in this comparison,
    for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process. Numbers in parentheses for the {\sc MG5\_aMC} simulations
    are rescaled to account for the effect related to the boson widths computed by {\sc MadSpin} (see the text for details).}
\end{table}

We now present the results of predictions matched to parton shower: the total rates within VBS cuts are displayed in Table~\ref{tab:PSrates}, both at LO and NLO
accuracy. For {\sc MG5\_aMC}, the numbers in parentheses are rescaled to take into account the width effects described in the above paragraph. Once this effect is taken into
account, total rates from different tools agree within few per cent. Larger discrepancies however will appear for differential observables, which we are going to discuss in
the following. For any observable, results will be presented in two plots, shown side-by-side. In the plot on the left (right), (N)LO+PS predictions are shown
with different colours in the main frame. In the inset, these predictions are compared in both cases with a fixed-order prediction at NLO accuracy (obtained with
{\sc VBFNLO} \emph{i.e.}\ the VBS approximation with $s$-channel contributions). For the differential observables, the {\sc MG5\_aMC} predictions are \emph{not} rescaled to compensate for the width effect mentioned above.

@@

In conclusion, the comparison of tools including matching with parton-shower clearly shows the benefits of the inclusion of NLO corrections: for most observables described
effectively at NLO accuracy differences between tools are at (or below) the $10\%$ level. {\bf MZ compare this number with scale uncertainties}
Some exceptions exist, \emph{e.g.}\ the rapidity separation of the two tagging jets, which
on the one hand clearly suggest not to rely on a single tool/parton shower, and on the other make it worth to investigate more in details the way QCD radiation is
generated {\bf MZ cite here H VBF at NNLO?}. Finally, the size of discrepancies for observables that are described at a lower perturbative accuracy, notably those related to the third jet, suggest that
experimental analyses should rely as least as possible on those observables and, in any case, use conservative estimate of the theory 
uncertainties. On the one hand, un order to improve the
description of these observables, a simulation of VBS+j at NLO accuracy, currently unavailable but within the reach of modern 
automated tools, is certainly desirable; on the other hand, measurements of processes with similar 
color flow (EW production of a single vector boson plus jets,
 VBF, \ldots) can certainly help in order to discriminate which tools perform better in the comparison with data.

>>>>

\begin{table}[h!]
    \centering
    \begin{tabular}{c|r@{ $\pm$ }l}
      Code  &  \multicolumn{2}{c}{$\sigma[\rm{fb}]$}  \\
        \hline\hline
        {\sc MG5\_aMC}+{\sc Pythia8}&  $1.352 $ & $0.003$  \\
        {\sc MG5\_aMC}+{\sc Herwig++}&  $1.343 $ & $ 0.003$  \\
        {\sc MG5\_aMC}+{\sc Pythia8}, $\Gamma_{\rm resc}$&  $1.275$ & $0.003$  \\
        {\sc MG5\_aMC}+{\sc Herwig++}, $\Gamma_{\rm resc}$&  $1.267$ & $ 0.003$  \\
        {\sc PHANTOM}+{\sc Pythia8} &  $1.235\phantom{0} $ & $0.001$  \\
        {\sc PHANTOM}+{\sc Herwig++} &  $1.260\phantom{0} $ & $0.001$  \\
        {\sc VBFNLO}+{\sc Herwig7} &  $1.3001$ & $0.0002$  \\
    \end{tabular}
    \caption{\label{tab:PSratesLO} Rates at LO-QCD accuracy matched to parton shower within VBS cuts obtained with the different codes used in this comparison,
    for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process. The {\sc MG5\_aMC} results with $\Gamma_{\rm resc}$ 
    are rescaled to account for the effect related to the boson widths computed by {\sc MadSpin} (see the text for details).}
\end{table}

\begin{table}[h!]
    \centering
    \begin{tabular}{c|r@{ $\pm$ }l}
      Code  &  \multicolumn{2}{c}{$\sigma[\rm{fb}]$}  \\
        \hline\hline
        {\sc MG5\_aMC}+{\sc Pythia8}&  $1.450 ^{+XX}_{-YY} {}^{+XX}_{-YY} $ & $0.004$  \\
        {\sc MG5\_aMC}+{\sc Herwig++}&  $1.445 $ & $0.004$  \\
        {\sc MG5\_aMC}+{\sc Pythia8}, $\Gamma_{\rm resc}$&  $1.368$ & $0.004$  \\
        {\sc MG5\_aMC}+{\sc Herwig++}, $\Gamma_{\rm resc}$&  $1.363$ & $0.004$  \\
        {\sc POWHEG}  &  $1.3642$ & $0.0004$  \\
        {\sc VBFNLO}+{\sc Herwig7-Dipole} &  $1.3389 ^{+0.0}_{-1.0}$ & $0.0006$  \\
        {\sc VBFNLO}+{\sc Herwig7-Default} &  $1.3067$ & $0.0006$  \\
    \end{tabular}
    \caption{\label{tab:PSratesNLO} Rates at NLO-QCD  accuracy matched to parton shower within VBS cuts obtained with the different codes used in this comparison,
    for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process. The {\sc MG5\_aMC} results with $\Gamma_{\rm resc}$ 
    are rescaled to account for the effect related to the boson widths computed by {\sc MadSpin} (see the text for details). For
    {\sc VBFNLO}+{\sc Herwig7-Dipole}, the three-point scale uncertainties is shown, while for  {\sc MG5\_aMC}+{\sc Pythia8} the two displayed uncertainties
are respectively the nine-point scale uncertainty and the PDF one. Theory uncertainties are in percent.}
\end{table}

We now present the results of predictions matched to parton shower: the total rates within VBS cuts are displayed in Tables~\ref{tab:PSratesLO} and
\ref{tab:PSratesNLO}, at LO and NLO
accuracy respectively. For {\sc MG5\_aMC}, 
the numbers whith $\Gamma_{\rm resc}$ are rescaled to 
take into account the width effects described in the above paragraph.d  At NLO accuracy, for {\sc MG5\_aMC} + {\sc Pythia8} and {\sc VBFNLO}, we also quote
theoretical uncertainties: for the former, we show both PDF and scale uncertainties, obtained via exact reweighting~\cite{Frederix:2011ss} by varying independently the renormalization and factorization
scale by a factor 2 around its central value, Eq.~\ref{eq:defscale} (nine-point variations); for the latter, we show the 
three-point scale uncertainties, obtained by considering correlated variations of the renormalization and factorization scales. 
We observe that, once the width effect is taken into
account, total rates from different tools agree within few per cent. Larger discrepancies however will appear for differential observables, which we are going to discuss in
the following. For any observable, results will be presented in two plots, shown side-by-side. In the plot on the left (right), (N)LO+PS predictions are shown
with different colours in the main frame. In the inset, these predictions are compared in both cases with a fixed-order prediction at NLO accuracy (obtained with
{\sc VBFNLO} \emph{i.e.}\ the VBS approximation with $s$-channel contributions). For the differential observables, the {\sc MG5\_aMC} predictions are \emph{not} rescaled to compensate for the width effect mentioned above. As for the table, we show theoretical uncertainties for the NLO+PS samples
obtained with {\sc VBFNLO} and {\sc MG5\_aMC}: 
again, for the first the band corresponds to three-point variations, while for the second the darker (lighter) band corresponds to nine-point 
scale variations (plus PDF uncertainties, linearly added).

@@

In conclusion, the comparison of tools including matching with parton-shower clearly shows the benefits of the inclusion of NLO corrections: for most observables described
effectively at NLO accuracy differences between tools are at (or below) the $10\%$ level. 
Some exceptions exist, \emph{e.g.}\ the rapidity separation of the two tagging jets, which
on the one hand clearly suggest not to rely on a single tool/parton shower, and on the other make it worth to investigate more in details the way QCD radiation is
generated {\bf MZ cite here H VBF at NNLO?}. It is remarkable the fact that, even for those observables that display small discrepancies,
the theoretical uncertainty estimated via scale variations systematically underestimates the spread of predictions. Again, this stresses the need
 to employ at least two different tools in order to have a conservative estimate of theoretical uncertainties. Finally, the size of discrepancies for observables that are described at a lower perturbative accuracy, notably those related to the third jet, suggest that
experimental analyses should rely as least as possible on those observables and, in any case, use conservative estimate of the theory 
uncertainties. On the one hand, un order to improve the
description of these observables, a simulation of VBS+j at NLO accuracy, currently unavailable but within the reach of modern 
automated tools, is certainly desirable; on the other hand, measurements of processes with similar 
color flow (EW production of a single vector boson plus jets,
 VBF, \ldots) can certainly help in order to discriminate which tools perform better in the comparison with data.
